\ttitle

### 2015-08-18 Chapters/BodyComp.tex
Major changes as below
- Thesis summary completed
- Many changes to thesis.cls and main.tex to format abstract page and university url, etc
- Update Intro-body comp - measuring body comp
- Zot.bin backed up and updated
- Conclusion chapter title renamed to Discussion
- It is 01:40 am - Have listened to Dr. Dre, Snoop Dogg and Eminem all day
- 2 shots of Jura and I am knackered yet sleepless after 2 strong coffees
Huge commit following on from Donny conversation
- Change all references to before period
- Remove post 2011/2012 references from introduction - one 2012 reference left was published

\section{Introduction}
Major abdominal surgery especially for pancreatic disease is associated with significant morbidity and mortality. Patient selection is as important as identifying surgical treatable pathology in ensuring optimal outcomes. \parencite{balthazar_acute_2002}

\subsection{Role of preoperative CPET}
The role of cardiopulmonary exercise testing in the preoperative evaluation and risk assessment/stratification of patients undergoing major thoracic and abdominal surgery has become well established. A number of studies have shown that poor aerobic fitness demonstrated by a low $\dot{V}_{O_2}$AT or low $\dot{V}_{O_2}$Peak or both as measured at cardiopulmonary exercise testing is associated with increased morbidity and mortality after major surgery including bariartic\parencite{mccullough_cardiorespiratory_2006}, pancreatic\parencite{chandrabalan_pre-operative_2013,ausania_effects_2012}, liver \parencite{epstein_aerobic_2004}, cardiothoracic\parencite{brunelli_risk_2010, campione_oxygen_2010,torchio_exercise_2010} and abdominal aortic aneurysm surgery.\parencite{carlisle_mid-term_2007,thompson_cardiopulmonary_2011} CPET is now routinely used as part of the preoperative processes used to select patients for surgery as well as to help in decision making regarding preoperative care including the need for additional tests, preoperative and intraoperative optimisation, admission to critical care and postoperative care.
Patients are sometimes denied surgery if their performance at cardiopulmonary exercise testing is felt to be poor based on currently available evidence.

@@

The most common parameters used to quantify perioperative risk in surgical patients are oxygen consumption at the anaerobic threshold ($\dot{V}_{O_2}$AT) and at peak exercise capacity ($\dot{V}_{O_2}$Peak). Conventionally these have been reported as per weight ratios (ml/kg/min) to allow comparison between patients. However, numerous studies on cardiorespiratory exercise physiology have reported that normalising $\dot{V}_{O_2}$ using total body weight leads to spurious correlation errors unfairly penalising obese subjects.\parencite{seltzer_body_1940, tanner_fallacy_1949, toth_examination_1993, batterham_modeling_1999, goran_total_2000, krachler_cardiopulmonary_2014}

@@

\subsection{Calculation of body composition}
Computed tomography of the abdomen, performed as part of the routine preoperative staging, was used to calculate body composition based on previously published and well established methods.\parencite{bredella_comparison_2010,shen_total_2004}

%http://regionstraumapro.com/post/16349545265 - Excellent description of CT window width(W) and level/center(C)
The coronal and saggital reconstructions were used to accurately identify the L3 and L4 vertebrae. The CT window width (W) was set at 400 and the level/center (C) was set at 40. This allowed tissue between -160 HU and +240 HU to be represented in grayscale with adequate contrast between tissues of interest. The cross-sectional images at these levels where then exported as gray-scale bitmap images. The scale in millimeters was included with every image. A representative image is shown in Fig. 1.The GNU Image Manipulation Program (GIMP), an advanced, free, open-source, raster graphics editor was used for analysis of all images (www.gimp.org). The use of GIMP to analyse cross-sectional imaging for body composition has been described previously although by using a different technique to what has been employed by us. \parencite{anblagan_measurement_2013}

@@

\subsection{Cardiopulmonary exercise testing}
All patients performed cardiopulmonary exercise testing on a cycle ergometer as described in Section \ref{sec:cpx_method}. Raw data of all breath-by-breath parameters averaged every 10 seconds was collected for analysis. The first three minutes of the recorded data were during the rest period when the patients were on the exercise bike but did not exercise. The average of each parameter measured between the first and second minute (6 readings) was treated as the rest value. Anaerobic threshold was identified using previously established methods and all corresponding parameters at this point were recorded.\parencite{beaver_new_1986,sue_metabolic_1988} Peak exercise was identified by the maximum oxygen consumption recorded towards the end of the exercise period and all other parameters recorded at this point were considered as peak exercise values.

@@

\subsection{Body composition and Clinico-pathological characteristics}
Eighty-two patients (35 male) were included in the study. The clinico-pathological characteristics of the study patients and their relationship to body composition is shown in Table \ref{table:bc_clinical} on page \pageref{table:bc_clinical}. There were several significant associations between clinico-pathological variables and body composition as depicted in this table.

	\input{Tables/c4_t1}

@@

Seltzer reported in his 1940 study of 34 subjects, that the individuals who were more "lateral" than "linear" had lower oxygen intakes per kilo body weight.\parencite{seltzer_body_1940} Tanner in his article titled \textit{"Fallacy of per-weight and per-surface area standards, and their relation to spurious correlation"}\parencite{tanner_fallacy_1949} in the Journal of Applied Physiology in 1947 recognised the dangers of expressing physiological variables as a function of total body mass. In a detailed anaysis comparing $\dot{V}_{O_2}$ and body build, he concludes that \textit{"as the index wt./stature increases, O2/wt. must be expected to decrease purely as a result of the method used for representing the data." }

Batterham et al studied 1314 apparently healthy men employed at the National Aeronautics and Space Administration Johnson Space Center in Houston, Texas.\parencite{batterham_modeling_1999} The authors report that as body mass increased, the proportion composed of fat-free mass decreased. They also found that fat-free mass had a linear relationship with oxygen consumption while total body mass did not. They suggest that ideally estimates of fat-free mass should be used in the representation of oxygen consumption to allow more reliable comparison between subjects.

Janz et al studied oxygen consumption and aerobic capacity in adolescents over several years as part of the Muscatine study and reported their findings in 1997\parencite{janz_three-year_1997} and 1998.\parencite{janz_longitudinal_1998} Aerobic capacity in the form of $\dot{V}_{O_2}$peak was evaluated annually in 126 children (mean age 10.3 years) for five years. Body composition changes were also tracked over this period. They reported on the changes in body composition that occur over time and the differences in these changes between circum-pubertal boys and girls. They reported on the significant difficulties in normalising $\dot{V}_{O_2}$ using total body mass and suggested that fat-free mass was the most appropriate variable for normalising $\dot{V}_{O_2}$. They found that $\dot{V}_{O_2}$ normalised using total body mass underestimated aerobic fitness levels of heavier boys and girls. However, this underestimation was greater in girls than in boys.

Goran et al reported that total body fat did not affect maximal aerobic capacity.\parencite{goran_total_2000} They reported on $\dot{V}_{O_2}$max in obese women before and after weight loss. $\dot{V}_{O_2}$max corrected for total body weight was significantly lower in the obese state while $\dot{V}_{O_2}$max corrected for fat-free mass did not change significantly after weight loss. They also reported that the limiting factor in the obese state was not the cardio-respiratory system but the fact that it was more difficult for obese individuals to do the same amount of work as a normal weight person in weight-bearing activities. This is likely due to the extra fat mass in these individuals that did not contribute to aerobic capacity but instead may increase the exercise load.

These findings have been replicated by several other authors in different subject groups.\parencite{loftin_scaling_2001,  lemaitre_maximum_2006,savonen_current_2012, krachler_cardiopulmonary_2014} Several of the above studies also recommend using allometric scaling to avoid the confounding effects of total body weight. However, this has not gained widespread clinical use.

In a study aimed at determining the optimal method of expressing $\dot{V}_{O_2}$max, Maciejczyk and coworkers analysed the differing influence of body fat and lean body mass on aerobic performance in a two groups of physically fit men categorised based on their body fat percentage.\parencite{maciejczyk_influence_2014} They reported that high body mass regardless of composition was correlated negatively with $\dot{V}_{O_2}$ when it was corrected for total body weight penalising otherwise fit men purely based on the proportion of body weight that was contributed by body fat. However, when $\dot{V}_{O_2}$ was corrected for lean body mass, they found that the results were similar between the low body fat and high fat body groups. They, similar to Goran et al\parencite{goran_total_2000}, recommend that $\dot{V}_{O_2}$ be normalised to lean body mass rather than total body weight.

@@

\subsection{Clinical implications of spurious correlation}
Older et al in their pioneering study in 1993 reported that $\dot{V}_{O_2}$AT < 11mls/kg/min was associated with increased mortality in elderly patients undergoing major abdominal surgery.\parencite{older_preoperative_1993} While they did not provide any data on other preoperative or intra-operative factors, they concluded that cardiopulmonary exercise testing was useful in predicting postoperative outcome. However, this first report on the use of cardiopulmonary exercise testing as a preoperative risk assessment tools repeatedly states that a $\dot{V}_{O_2}$AT < 11mls/kg/min represented cardiac failure. This association is repeated in their later work on 548 patients which also showed a clear association between $\dot{V}_{O_2}$AT < 11mls/kg/min and mortality due to cardiovascular causes. \parencite{older_cardiopulmonary_1999} The concepts of \textit{'surgical anaerobic threshold'} and \textit{'postoperative cardiac failure'} were introduced later and were described as the \textit{'inability of the heart to meet the demand of postoperative stress.'}\parencite{society_ats/accp_2003}

Swart and Carlisle reported that $\dot{V}_{O_2}$AT < 11mls/kg/min in patients undergoing open colorectal surgery was associated with adverse outcomes.\parencite{swart_case-controlled_2012} However, the proportion of females in the low $\dot{V}_{O_2}$AT group was significantly greater than that in the normal $\dot{V}_{O_2}$AT groups (24\% vs 51\%). The average $\dot{V}_{O_2}$AT in men calculated from the data presented in their paper was 11.02 mls/kg/min while in women it was 9.81 mls/kg/min. In a study by Wilson et al that reported cardiopulmonary exercise testing predicted outcome in major elective intra-abdominal surgery, the proportion of females in the low $\dot{V}_{O_2}$AT group was 51\% while it was 28\% in the group with normal AT.\parencite{wilson_impaired_2010} There was no data presented on body mass index in this study.

@@

Where time to surgery is not critical, prehabilitation has gained an increasingly important role in optimising patients for surgery and mitigating the effects of neoadjuvant oncological therapy. Cardiopulmonary exercise testing has been reported to be a useful objective measure of the impact of prehabilitation in surgical patients.\parencite{west_effect_2015}

>>>>

\section{Introduction}
Major abdominal surgery especially for pancreatic disease is associated with significant morbidity and mortality. Patient selection is as important as identifying surgical treatable pathology in ensuring optimal outcomes \parencite{balthazar_acute_2002}.

\subsection{Role of preoperative CPET}
The role of cardiopulmonary exercise testing in the preoperative evaluation and risk assessment/stratification of patients undergoing major thoracic and abdominal surgery has become well established. A number of studies have shown that poor aerobic fitness demonstrated by a low $\dot{V}_{O_2}$AT or low $\dot{V}_{O_2}$Peak or both as measured at cardiopulmonary exercise testing is associated with increased morbidity and mortality after major surgery including bariartic\parencite{mccullough_cardiorespiratory_2006}, pancreatic\parencite{ausania_effects_2012}[Chapter \ref{ch_cpet_outcomes}], liver \parencite{epstein_aerobic_2004}, cardiothoracic\parencite{brunelli_risk_2010, campione_oxygen_2010,torchio_exercise_2010} and abdominal aortic aneurysm surgery \parencite{carlisle_mid-term_2007,thompson_cardiopulmonary_2011}. CPET is now routinely used as part of the preoperative processes used to select patients for surgery as well as to help in decision making regarding preoperative care including the need for additional tests, preoperative and intraoperative optimisation, admission to critical care and postoperative care.
Patients are sometimes denied surgery if their performance at cardiopulmonary exercise testing is felt to be poor based on currently available evidence.

@@

The most common parameters used to quantify perioperative risk in surgical patients are oxygen consumption at the anaerobic threshold ($\dot{V}_{O_2}$AT) and at peak exercise capacity ($\dot{V}_{O_2}$Peak). Conventionally these have been reported as per weight ratios (ml/kg/min) to allow comparison between patients. However, numerous studies on cardiorespiratory exercise physiology have reported that normalising $\dot{V}_{O_2}$ using total body weight leads to spurious correlation errors unfairly penalising obese subjects \parencite{seltzer_body_1940, tanner_fallacy_1949, toth_examination_1993, batterham_modeling_1999, goran_total_2000, krachler_cardiopulmonary_2014}

@@

\subsection{Calculation of body composition}
Computed tomography of the abdomen, performed as part of the routine preoperative staging, was used to calculate body composition based on previously published and well established methods \parencite{bredella_comparison_2010,shen_total_2004}.

%http://regionstraumapro.com/post/16349545265 - Excellent description of CT window width(W) and level/center(C)
The coronal and sagittal reconstructions were used to accurately identify the L3 and L4 vertebrae. The CT window width (W) was set at 400 and the level/center (C) was set at 40. This allowed tissue between -160 HU and +240 HU to be represented in grayscale with adequate contrast between tissues of interest. The cross-sectional images at these levels where then exported as gray-scale bitmap images. The scale in millimeters was included with every image. A representative image is shown in Fig. 1.The GNU Image Manipulation Program (GIMP), an advanced, free, open-source, raster graphics editor was used for analysis of all images (www.gimp.org). The use of GIMP to analyse cross-sectional imaging for body composition has been described previously although by using a different technique to what has been employed by us \parencite{anblagan_measurement_2013}.

@@

\subsection{Cardiopulmonary exercise testing}
All patients performed cardiopulmonary exercise testing on a cycle ergometer as described in Section \ref{sec:cpx_method}. Raw data of all breath-by-breath parameters averaged every 10 seconds was collected for analysis. The first three minutes of the recorded data were during the rest period when the patients were on the exercise bike but did not exercise. The average of each parameter measured between the first and second minute (6 readings) was treated as the rest value. Anaerobic threshold was identified using previously established methods and all corresponding parameters at this point were recorded \parencite{beaver_new_1986,sue_metabolic_1988}. Peak exercise was identified by the maximum oxygen consumption recorded towards the end of the exercise period and all other parameters recorded at this point were considered as peak exercise values.

@@

\subsection{Body composition and Clinico-pathological characteristics}
Eighty-two patients (35 male) were included in the study. The clinico-pathological characteristics of the study patients and their relationship to body composition is shown in Table \ref{table:bc_clinical} on page \pageref{table:bc_clinical}. There were several significant associations between clinico-pathological variables and body composition as depicted in this table.
	\input{Tables/bc_sex_specific_distrib}
	\input{Tables/bc_clinical}
	\input{Tables/c4_t1}

@@

Seltzer reported in his 1940 study of 34 subjects, that the individuals who were more "lateral" than "linear" had lower oxygen intakes per kilo body weight

Batterham et al studied 1314 apparently healthy men employed at the National Aeronautics and Space Administration Johnson Space Center in Houston, Texas \parencite{batterham_modeling_1999}. The authors report that as body mass increased, the proportion composed of fat-free mass decreased. They also found that fat-free mass had a linear relationship with oxygen consumption while total body mass did not. They suggest that ideally estimates of fat-free mass should be used in the representation of oxygen consumption to allow more reliable comparison between subjects.

Janz et al studied oxygen consumption and aerobic capacity in adolescents over several years as part of the Muscatine study and reported their findings in 1997\parencite{janz_three-year_1997} and 1998 \parencite{janz_longitudinal_1998}. Aerobic capacity in the form of $\dot{V}_{O_2}$peak was evaluated annually in 126 children (mean age 10.3 years) for five years. Body composition changes were also tracked over this period. They reported on the changes in body composition that occur over time and the differences in these changes between circum-pubertal boys and girls. They reported on the significant difficulties in normalising $\dot{V}_{O_2}$ using total body mass and suggested that fat-free mass was the most appropriate variable for normalising $\dot{V}_{O_2}$. They found that $\dot{V}_{O_2}$ normalised using total body mass underestimated aerobic fitness levels of heavier boys and girls. However, this underestimation was greater in girls than in boys.

Goran et al reported that total body fat did not affect maximal aerobic capacity \parencite{goran_total_2000}. They reported on $\dot{V}_{O_2}$max in obese women before and after weight loss. $\dot{V}_{O_2}$max corrected for total body weight was significantly lower in the obese state while $\dot{V}_{O_2}$max corrected for fat-free mass did not change significantly after weight loss. They also reported that the limiting factor in the obese state was not the cardio-respiratory system but the fact that it was more difficult for obese individuals to do the same amount of work as a normal weight person in weight-bearing activities. This is likely due to the extra fat mass in these individuals that did not contribute to aerobic capacity but instead may increase the exercise load.

These findings have been replicated by several other authors in different subject groups \parencite{loftin_scaling_2001,  lemaitre_maximum_2006,savonen_current_2012, krachler_cardiopulmonary_2014}. Several of the above studies also recommend using allometric scaling to avoid the confounding effects of total body weight. However, this has not gained widespread clinical use.

In a study aimed at determining the optimal method of expressing $\dot{V}_{O_2}$max, Maciejczyk and coworkers analysed the differing influence of body fat and lean body mass on aerobic performance in a two groups of physically fit men categorised based on their body fat percentage \parencite{maciejczyk_influence_2014}. They reported that high body mass regardless of composition was correlated negatively with $\dot{V}_{O_2}$ when it was corrected for total body weight penalising otherwise fit men purely based on the proportion of body weight that was contributed by body fat. However, when $\dot{V}_{O_2}$ was corrected for lean body mass, they found that the results were similar between the low body fat and high fat body groups. They, similar to Goran et al\parencite{goran_total_2000}, recommend that $\dot{V}_{O_2}$ be normalised to lean body mass rather than total body weight.

@@

\subsection{Clinical implications of spurious correlation}
Older et al in their pioneering study in 1993 reported that $\dot{V}_{O_2}$AT < 11mls/kg/min was associated with increased mortality in elderly patients undergoing major abdominal surgery \parencite{older_preoperative_1993}. While they did not provide any data on other preoperative or intra-operative factors, they concluded that cardiopulmonary exercise testing was useful in predicting postoperative outcome. However, this first report on the use of cardiopulmonary exercise testing as a preoperative risk assessment tools repeatedly states that a $\dot{V}_{O_2}$AT < 11mls/kg/min represented cardiac failure. This association is repeated in their later work on 548 patients which also showed a clear association between $\dot{V}_{O_2}$AT < 11mls/kg/min and mortality due to cardiovascular causes \parencite{older_cardiopulmonary_1999}. The concepts of \textit{'surgical anaerobic threshold'} and \textit{'postoperative cardiac failure'} were introduced later and were described as the \textit{'inability of the heart to meet the demand of postoperative stress.'}\parencite{society_ats/accp_2003}.

Swart and Carlisle reported that $\dot{V}_{O_2}$AT < 11mls/kg/min in patients undergoing open colorectal surgery was associated with adverse outcomes \parencite{swart_case-controlled_2012}. However, the proportion of females in the low $\dot{V}_{O_2}$AT group was significantly greater than that in the normal $\dot{V}_{O_2}$AT groups (24\% vs 51\%). The average $\dot{V}_{O_2}$AT in men calculated from the data presented in their paper was 11.02 mls/kg/min while in women it was 9.81 mls/kg/min. In a study by Wilson et al that reported cardiopulmonary exercise testing predicted outcome in major elective intra-abdominal surgery, the proportion of females in the low $\dot{V}_{O_2}$AT group was 51\% while it was 28\% in the group with normal AT \parencite{wilson_impaired_2010}. There was no data presented on body mass index in this study.

@@

Where time to surgery is not critical, prehabilitation has gained an increasingly important role in optimising patients for surgery and mitigating the effects of neoadjuvant oncological therapy. Cardiopulmonary exercise testing has been reported to be a useful objective measure of the impact of prehabilitation in surgical patients \parencite{west_effect_2015}.

### 2015-08-18 Chapters/CPET_Jaundice.tex
Huge commit following on from Donny conversation
- Change all references to before period
- Remove post 2011/2012 references from introduction - one 2012 reference left was published online in Aug 2011
- Rename title - 'investigation into clinical utility of CPET'
- Remove hyphenation of words
- Left justify with ragged right edge
- Remove bold formatting in some tables
- Minor grammar and style corrections in multiple places
- Change reference Chandrabalan 2013 to Chapter 2
- Create some new tables for body comp chapter based on Colin Richards paper - these dont work

\section{Introduction}
Patients with tumours involving the pancreatic head or the periampullary region often present with inoperable disease. In the minority of patients with operable disease, resectional surgery in the form of a pancreaticoduodenectomy remains the main modality of treatment and only chance of a potential cure. However, major pancreatic surgery is associated with significant morbidity and mortality and is only undertaken in specialist centres. Patient selection, preoperative optimisation, good surgical technique and improvements in postoperative care have all contributed to a reduction in mortality\parencite{winter_1423_2006} but morbidity remains high. While several technical strategies have been described in recent years to minimise morbidity, these strategies are not necessarily based on a better understanding of the physiological basis of postoperative complications in these patients.

@@

Obstructive jaundice has been reported to be associated with abnormal cardiovascular physiology in several animal and human studies. Surgery in the jaundiced patient has been reported to be associated with adverse postoperative haemodynamic events and renal dysfunction.\parencite{pain_perioperative_1985,green_systemic_1995} The association between jaundice and cardiovascular physiology was reported over a hundred years ago by King and co-workers who found that injection of porcine bile pigment into dogs resulted in bradycardia, hypotension and eventually death.\parencite{king_effect_1909} Green and co-workers (1986) described the effects of ‘cholemia' in dogs that were subjected to choledochocaval anastomosis. The resultant myocardial depression was described by them as the ‘jaundiced heart'\parencite{green_jaundiced_1986} and has been reported to be associated with poor myocardial response to inotropic stimulation in dogs\parencite{binah_obstructive_1985, bomzon_systemic_1986} as well as humans.\parencite{lumlertgul_jaundiced_1991}

@@

\subsection{Cardiopulmonary Exercise Test}
Cardiopulmonary exercise tests were performed in the Department of Respiratory Medicine at the Glasgow Royal Infirmary using the ZAN-600 CPET suite (nSpire Health, Longmont, CO 80501, USA) (9). All patients underwent standard pulmonary function tests and spirometry prior to cardiopulmonary exercise testing. A cycle ergometer was used to perform a symptom-limited, incremental work-load test preceded by a 3-minute rest period. The test was stopped when patients achieved their maximum exercise tolerance, when significant ischaemic changes occurred on ECG or for other safety reasons. Peak oxygen consumption achieved at this stage was defined as $\dot{V}_{O_2}$Peak. The $\dot{V}_{O_2}$AT was calculated using the V-slope\parencite{beaver_new_1986,sue_metabolic_1988} and ventilatory equivalents\parencite{society_ats/accp_2003} methods. $\dot{V}_{O_2}$AT less than 10 ml/kg/min was considered to be low based on previous work by us\parencite{chandrabalan_pre-operative_2013} as well as Ausania and co-workers\parencite{ausania_effects_2012} which has shown increased incidence of complications in these patients. Oxygen consumption at peak exercise ($\dot{V}_{O_2}$Peak) was dichotomised using a cut-off of 16 ml/kg/min. Detailed description of cardiopulmonary exercise testing as well as the physiological parameters described in this study are published elsewhere.\parencite{balady_clinicians_2010}

@@

The use of CPET in preoperative risk prediction was first made popular over two decades ago by Older and co-workers.\parencite{older_preoperative_1993} Since then cardiopulmonary exercise testing has been reported to be useful in identifying high risk patients prior to major general\parencite{snowden_submaximal_2010}, pancreatic\parencite{chandrabalan_pre-operative_2013,ausania_effects_2012}, oesophagogastric\parencite{nagamatsu_preoperative_2001} as well as vascular\parencite{carlisle_mid-term_2007} surgery. Cardiopulmonary exercise testing has been reported to be superior to conventional measures of comorbidity chiefly due to the dynamic nature of the test that evaluates the adequacy of oxygen delivery to tissues under physiological stress. However, the factors responsible for poor aerobic capacity in preoperative patients have not been adequately studied.

The association between jaundice and cardiovascular physiology was reported over a hundred years ago by King and co-workers who found that injection of porcine bile pigment into dogs resulted in bradycardia, hypotension and eventually death.\parencite{king_effect_1909}

Jaundice has been reported to be associated with myocardial depression\parencite{green_jaundiced_1986}, poor myocardial response to inotropic stimulation\parencite{lumlertgul_jaundiced_1991}, impaired sympathetic baroreflex sensitivity\parencite{song_baroreflex_2009}, deranged atrial natriuretic peptide levels\parencite{pereira_increased_1994,gallardo_increased_1998} as well as multiple other bile-acid receptor mediated effects on the cardiovascular system.\parencite{khurana_bile_2011} Moreover, some of these effects appear to be partly reversible by biliary drainage as demonstrated by Padillo and coworkers.\parencite{padillo_improved_2001}

@@

Pitt and coworkers in a prospective randomised trial comparing outcomes in jaundiced patients undergoing surgery with or without PBD reported that PBD was associated with increased cost without any decrease in postoperative complications.\parencite{pitt_does_1985} But, this study looked at a heterogenous group of patients of which only 7 underwent pancreaticoduodenectomy.

A recent meta-analysis\parencite{sewnath_meta-analysis_2002} analysed data from 5 randomised controlled trials comparing surgery with PBD versus surgery without PBD and concluded that PBD not only did not improve postoperative complication rates or mortality but resulted in a higher overall complication rate due to the morbidity associated with the procedure itself. All five RCTs included in this meta-analysis included a heterogenous group of operations with only a few undergoing pancreaticoduodenectomy while more than 50\% of patients underwent palliative bypass or exploratory laparotomy making comparison of outcomes difficult. A recent Cochrane Collaboration review of six trials including 520 patients concluded that PBD may be associated with serious adverse events and must not be performed routinely outwith trial settings.\parencite{wang_preoperative_2008}

The DROP trial sought to clarify the role of PBD in patients undergoing pancreaticoduodenectomy.\parencite{van_der_gaag_preoperative_2010} It randomised patients with bilirubin levels between 40 and 250 either to undergo surgery without PBD or to undergo PBD followed by surgery after 4 - 6 weeks. The authors reported that PBD resulted in an increase in incidence of complications of which the majority were related to the drainage procedure itself. However, this trial excluded patients with bilirubin levels over 250.

@@

However, such an association has been previously reported.\parencite{horwich_relationship_2009}  This may reflect the difficulty in obtaining accurate $\dot{V}_{O_2}$AT values in obese patients as a result of the calculations involved rather than due to true cardiopulmonary dysfunction. Other authors have suggested that different thresholds for CPET parameters may have to be considered in obese patients to improve risk-prediction.\parencite{donnelly_criteria_1990,hulens_exercise_2001} Cardiopulmonary exercise testing measures oxygen delivery to skeletal muscle. Adipose tissue, however, does not contribute to the metabolic activity that is measured during CPET. However, AT as normally reported, is calculated by dividing the oxygen consumption per minute at the ‘anaerobic threshold' into the weight of the patient. However, this does not account for the disproportionately higher amount of adipose tissue in overweight/obese patients resulting in a spuriously low AT (in mls/kg/min). The present study found no association between cardiorespiratory comorbidity and $\dot{V}_{O_2}$AT. Low $\dot{V}_{O_2}$AT in female patients and overweight/obese patients should be interpreted with caution as this may not be due to true poor aerobic capacity.

>>>>

\section{Introduction}
Patients with tumours involving the pancreatic head or the periampullary region often present with inoperable disease. In the minority of patients with operable disease, resectional surgery in the form of a pancreaticoduodenectomy remains the main modality of treatment and only chance of a potential cure. However, major pancreatic surgery is associated with significant morbidity and mortality and is only undertaken in specialist centres. Patient selection, preoperative optimisation, good surgical technique and improvements in postoperative care have all contributed to a reduction in mortality \parencite{winter_1423_2006} but morbidity remains high. While several technical strategies have been described in recent years to minimise morbidity, these strategies are not necessarily based on a better understanding of the physiological basis of postoperative complications in these patients.

@@

Obstructive jaundice has been reported to be associated with abnormal cardiovascular physiology in several animal and human studies. Surgery in the jaundiced patient has been reported to be associated with adverse postoperative haemodynamic events and renal dysfunction \parencite{pain_perioperative_1985,green_systemic_1995}. The association between jaundice and cardiovascular physiology was reported over a hundred years ago by King and co-workers who found that injection of porcine bile pigment into dogs resulted in bradycardia, hypotension and eventually death \parencite{king_effect_1909}. Green and co-workers (1986) described the effects of ‘cholemia' in dogs that were subjected to choledochocaval anastomosis. The resultant myocardial depression was described by them as the ‘jaundiced heart'\parencite{green_jaundiced_1986} and has been reported to be associated with poor myocardial response to inotropic stimulation in dogs \parencite{binah_obstructive_1985, bomzon_systemic_1986} as well as humans \parencite{lumlertgul_jaundiced_1991}.

@@

\subsection{Cardiopulmonary Exercise Test}
Cardiopulmonary exercise tests were performed in the Department of Respiratory Medicine at the Glasgow Royal Infirmary using the ZAN-600 CPET suite (nSpire Health, Longmont, CO 80501, USA) (9). All patients underwent standard pulmonary function tests and spirometry prior to cardiopulmonary exercise testing. A cycle ergometer was used to perform a symptom-limited, incremental work-load test preceded by a 3-minute rest period. The test was stopped when patients achieved their maximum exercise tolerance, when significant ischaemic changes occurred on ECG or for other safety reasons. Peak oxygen consumption achieved at this stage was defined as $\dot{V}_{O_2}$Peak. The $\dot{V}_{O_2}$AT was calculated using the V-slope \parencite{beaver_new_1986,sue_metabolic_1988} and ventilatory equivalents \parencite{society_ats/accp_2003} methods. $\dot{V}_{O_2}$AT less than 10 ml/kg/min was considered to be low based on previous work by us (Chapter \ref{ch_cpet_outcomes}) as well as Ausania and co-workers \parencite{ausania_effects_2012} which has shown increased incidence of complications in these patients. Oxygen consumption at peak exercise ($\dot{V}_{O_2}$Peak) was dichotomised using a cut-off of 16 ml/kg/min. Detailed description of cardiopulmonary exercise testing as well as the physiological parameters described in this study are published elsewhere \parencite{balady_clinicians_2010}.

@@

The use of CPET in preoperative risk prediction was first made popular over two decades ago by Older and co-workers \parencite{older_preoperative_1993}. Since then cardiopulmonary exercise testing has been reported to be useful in identifying high risk patients prior to major general \parencite{snowden_submaximal_2010}, pancreatic \parencite{ausania_effects_2012},[Chapter \ref{ch_cpet_outcomes}], oesophagogastric \parencite{nagamatsu_preoperative_2001} as well as vascular \parencite{carlisle_mid-term_2007} surgery. Cardiopulmonary exercise testing has been reported to be superior to conventional measures of comorbidity chiefly due to the dynamic nature of the test that evaluates the adequacy of oxygen delivery to tissues under physiological stress. However, the factors responsible for poor aerobic capacity in preoperative patients have not been adequately studied.

The association between jaundice and cardiovascular physiology was reported over a hundred years ago by King and co-workers who found that injection of porcine bile pigment into dogs resulted in bradycardia, hypotension and eventually death \parencite{king_effect_1909}.

Jaundice has been reported to be associated with myocardial depression \parencite{green_jaundiced_1986}, poor myocardial response to inotropic stimulation \parencite{lumlertgul_jaundiced_1991}, impaired sympathetic baroreflex sensitivity \parencite{song_baroreflex_2009}, deranged atrial natriuretic peptide levels \parencite{pereira_increased_1994,gallardo_increased_1998} as well as multiple other bile-acid receptor mediated effects on the cardiovascular system \parencite{khurana_bile_2011}. Moreover, some of these effects appear to be partly reversible by biliary drainage as demonstrated by Padillo and coworkers \parencite{padillo_improved_2001}.

@@

Pitt and coworkers in a prospective randomised trial comparing outcomes in jaundiced patients undergoing surgery with or without PBD reported that PBD was associated with increased cost without any decrease in postoperative complications \parencite{pitt_does_1985}. But, this study looked at a heterogenous group of patients of which only 7 underwent pancreaticoduodenectomy.

A recent meta-analysis \parencite{sewnath_meta-analysis_2002} analysed data from 5 randomised controlled trials comparing surgery with PBD versus surgery without PBD and concluded that PBD not only did not improve postoperative complication rates or mortality but resulted in a higher overall complication rate due to the morbidity associated with the procedure itself. All five RCTs included in this meta-analysis included a heterogenous group of operations with only a few undergoing pancreaticoduodenectomy while more than 50\% of patients underwent palliative bypass or exploratory laparotomy making comparison of outcomes difficult. A recent Cochrane Collaboration review of six trials including 520 patients concluded that PBD may be associated with serious adverse events and must not be performed routinely outwith trial settings \parencite{wang_preoperative_2008}.

The DROP trial sought to clarify the role of PBD in patients undergoing pancreaticoduodenectomy \parencite{van_der_gaag_preoperative_2010}. It randomised patients with bilirubin levels between 40 and 250 either to undergo surgery without PBD or to undergo PBD followed by surgery after 4 - 6 weeks. The authors reported that PBD resulted in an increase in incidence of complications of which the majority were related to the drainage procedure itself. However, this trial excluded patients with bilirubin levels over 250.

@@

However, such an association has been previously reported \parencite{horwich_relationship_2009}  This may reflect the difficulty in obtaining accurate $\dot{V}_{O_2}$AT values in obese patients as a result of the calculations involved rather than due to true cardiopulmonary dysfunction. Other authors have suggested that different thresholds for CPET parameters may have to be considered in obese patients to improve risk-prediction \parencite{donnelly_criteria_1990,hulens_exercise_2001}. Cardiopulmonary exercise testing measures oxygen delivery to skeletal muscle. Adipose tissue, however, does not contribute to the metabolic activity that is measured during CPET. However, AT as normally reported, is calculated by dividing the oxygen consumption per minute at the ‘anaerobic threshold' into the weight of the patient. However, this does not account for the disproportionately higher amount of adipose tissue in overweight/obese patients resulting in a spuriously low AT (in mls/kg/min). The present study found no association between cardiorespiratory comorbidity and $\dot{V}_{O_2}$AT. Low $\dot{V}_{O_2}$AT in female patients and overweight/obese patients should be interpreted with caution as this may not be due to true poor aerobic capacity.
